A GPU Implementation of Capon for an Active Sonar System

### 2012-05-07 ECUA2012/gpu_capon_ecua2012.tex
la2pdf commit - 2012:05:07	11:52:36	(Mon 7.May)

We have tested \gls{GPU} implementation of the Capon beamformer on experimental data from 32 element Kongsberg Maritime HISAS1030 synthetic aperture sonar~\cite{Hansen2009}. The sonar was attached to the HUGIN \gls{AUV} and used in strip-map mode, and the imaged object was the 1500 dwt oil tanker wreck Holmengraa (Figure \ref{holmengraa}). Holmengraa is roughly 68\,m long and 9\,m wide, and lies on a slanted seabed at 77\,m depth. The Capon image was created with the parameters $L=16$, $N_\text{avg}=3$, and $d=1\%$, which proved to be a reasonable selection for this scenario.

\begin{figure}[!t]
\centering
\includegraphics[width=\linewidth]{gfx/benchmark.pdf}
\caption{Capon benchmarks.}\label{benchmarks}
\end{figure}

To test the runtime speed of our \gls{GPU} implementation of Capon, we have compared it to an optimized single thread $C$ and $Matlab$ implementation (Figure \ref{benchmarks}). The test system was a quad-core Intel Xeon E5620, with 64Gb of \gls{RAM}, and an nVidia Quadro 6000 card. This is a \gls{GPU} with 448 \gls{CUDA} cores and 6Gb of onboard \gls{RAM}, capable of roughly 1Tflops. We kept the channel count at $M=32$, and processed a $640\times 86$ pixel image with $d=2\%$ diagonal loading.

@@

Figure \ref{benchmarks} shows how a \gls{GPU} can be used to improve matters. The test image here was 55kpx, hence much smaller than the Holmengraa image, but the relative performance increase seen when accelerating the algorithm with a \gls{GPU} is still accurate.

>>>>

We have tested our \gls{GPU} implementation of the Capon beamformer on experimental data from the 32 element Kongsberg Maritime HISAS1030 synthetic aperture sonar~\cite{Hansen2009}. The sonar was attached to the HUGIN \gls{AUV} and used in strip-map mode, and the imaged object was the 1500 dwt oil tanker wreck Holmengraa (Figure \ref{holmengraa}). Holmengraa is roughly 68\,m long and 9\,m wide, and lies on a slanted seabed at 77\,m depth. The Capon image was created with the parameters $L=16$, $N_\text{avg}=3$, and $d=1\%$, which proved to be a reasonable selection for this scenario.

\begin{figure}[!t]
\centering
\includegraphics[width=\linewidth]{gfx/benchmark.pdf}
\caption{Capon benchmarks. An 55\,kpx image processed with $M=32$ channels and $d=2\%$ diagonal loading. }\label{benchmarks}
\end{figure}

To test the runtime speed of our \gls{GPU} implementation of Capon, we have compared it to an optimized single thread C and Matlab implementation (Figure \ref{benchmarks}). The test system was a quad-core Intel Xeon E5620, with 64Gb of \gls{RAM}, and an nVidia Quadro 6000 card. This is a \gls{GPU} with 448 \gls{CUDA} cores and 6Gb of onboard \gls{RAM}, capable of roughly 1Tflops. We kept the channel count at $M=32$, and processed a 55\,kpixel image with $d=2\%$ diagonal loading.

@@

Figure \ref{benchmarks} shows how a \gls{GPU} can be used to improve matters.  The test image used here was 55kpx,   and the \gls{} hence the runtimes are seen are   much smaller than the Holmengraa image, but the relative performance increase seen when accelerating the algorithm with a \gls{GPU} is still accurate.
